\tname{}: Intrusion Detection Rule Synthesis

### 2020-06-06 paper/paper.tex
minor changes
merging

%\title{\tname{} for NIDS Rule Synthesis}
\title{Rule Synthesis for Network Intrusion Detectors}

@@

\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.
\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
we call \suri\ directly, through a public API, by providing a
\emph{pcap} files on input. \suri\ is invoked multiple times for
creating rules.

@@

\noindent
\textbf{Alternative methods for rule synthesis.} There are important
differences between our work and prior work. Vollmer
\etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
derive rules for Snort, which are very similar to \suri's.  Their
solution focused on icmp attacks for which plausible attacks can be
created directly from packets (see
Section~\ref{sec:rules-and-packets}). They modeled every field of the
icmp packet in chromosomes of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modeling the packet fields. They
showed that they were able to generate rules without false negatives
and only 3 false positives across 33K benign packets. The key
differences between our work and Vollmer~\etal{}'s are 1) the support
of different protocols and 2) the use of ranking to discriminate
alternative plausible rules. As for 1, our approach supports a variety
of protocols and options (\eg{}, options that refer to message
payloads). The icmp rules they generate only refer to message fields.
\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a
    secao 2.3 diz q \textbf{geralmente} as regras nao-http possuem
    apenas field options}. } As for 2, our solution additionally ranks
plausible options, which contain \emph{no} false positives or
negatives with respect to the benign and malicious traffic
considered. Guruprasad~\etal{}~\cite{guruprasad} also used an
evolutionary approach to solve the problem. Although the solution of
Guruprasad~\etal{} supports more protocols than Vollmer~\etal{}', our
approach still supports a larger set of protocols and the ability to
ranks rules. Muthuregunathan~\etal{}~\cite{muthuregunathan-etal2009}
propose a clustering method to create rules. They characterize
candidate solutions as a set of field values and use a clustering
algorithm---hybrid k-means partitioned around menoids---to isolate
malicious from benign traffic. Subsequently, they optimize the formed
clusters using a genetic algorithm and hill-climbing search. Ou
approach differs from theirs in the method used, results, and support
for content options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed
ARGS, a system to synthesize http rules that take malicious traffic as
input. ARGS looks for a common pattern across the different inputs to
define rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic. Fallahi
\etal{}\cite{fallahi} proposed flow-based synthesis approach to rule
generation. Instead of analyzing packets individually, their algorithm
investigates and creates the rules for flows in the network. The
algorithm was tested using the ISCX 2012 dataset and showed good
results in terms of precision, but it is limited to that type of
attack. As it won't analyse individual details of each packet, it
wouldn't be able to identify and generate rules for attacks that
require detailed options, such as the HTTP PESC, RCE and LFI.

>>>>

%\title{\tname{} for NIDS Rule Synthesis}
\title{\tname{}: Intrusion Detection Rule Synthesis}  
%\title{Rule Synthesis for Network Intrusion Detectors}

@@

\noindent\textbf{Correctness argument.}~The algorithm reports rules that
capture only the positive traffic. Every rule created during the
search captures the attack associated with the input as every rule
created contains a proper subset of the options from the original rule
and all those options are, by definition, satisfied by the traffic
produced by the original attack. In addition, from the definition of
$\mathit{isPlausible}$, reported rules capture \emph{only} the
positive traffic. If negative traffic is captured, a rule is not added
to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
O$ and reported on the output. Finally, from the definition of
function $\mathit{filter}$, only the rules in $\mathcal O$ that
capture all the instances of the attack are reported on output.

\noindent\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
we call \suri\ directly, through a public API, by providing a
\emph{pcap} files on input. \suri\ is invoked multiple times for
creating rules.

@@

% There are important differences between our work and prior work.

\noindent
\textbf{Alternative methods for rule synthesis.}~Vollmer
\etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
derive rules for Snort, which are very similar to \suri's.  Their
solution focused on icmp attacks for which plausible attacks can be
created directly from packets (see
Section~\ref{sec:rules-and-packets}). They modeled every field of the
icmp packet in chromosomes of candidate solutions (\ie{}, a rule) of a
genetic search. The initial population is a set of candidate solutions
that randomly combine options, modelling the packet fields. They were
able to generate rules without false negatives and only 3 false
positives across 33K benign packets. The key differences between our
work and Vollmer~\etal{}'s are 1) the support of different protocols
and 2) the use of ranking to discriminate alternative plausible
rules. As for 1, our approach supports a variety of protocols and
options (\eg{}, options that refer to message payloads). The icmp
rules they generate only refer to message fields.  \MyComment{\Luc{eh
    possivel incluir contents em regras icmp tb. o q a secao 2.3 diz q
    \textbf{geralmente} as regras nao-http possuem apenas field
    options}. } As for 2, our solution additionally ranks plausible
options, which contain \emph{no} false positives or negatives with
respect to the benign and mallicious traffic
considered. Guruprasad~\etal{}~\cite{guruprasad} also used an
evolutionary approach to solve the problem. Although the solution of
Guruprasad~\etal{} supports more protocols than Vollmer~\etal{}', our
approach still supports a larger set of protocols and the ability to
ranks rules. Muthuregunathan~\etal{}~\cite{muthuregunathan-etal2009}
propose a clustering method to create rules. They characterize
candidate solutions as a set of field values and use a clustering
algorithm---hybrid k-means partitioned around menoids---to isolate
malicious from benign traffic. Subsequently, they optimize the formed
clusters using a genetic algorithm and hill-climbing search. Ou
approach differs from theirs in the method used, results, and support
for content options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed
ARGS, a system to synthesize http rules that take malicious traffic as
input. ARGS looks for a common pattern across the different inputs to
define rules. It does not take existing rules as input or benign
traffic. Consequently, false positive can be produced and need to be
handled manually by users. Our approach, in contrast, does not
generate false positives w.r.t. the input traffic. Fallahi
\etal{}\cite{fallahi} proposed flow-based synthesis approach to rule
generation. Instead of analyzing packets individually, their algorithm
investigates and creates the rules for flows in the network. The
algorithm was tested using the ISCX 2012 dataset and showed good
results in terms of precision, but it is limited to that type of
attack. As it won't analyse individual details of each packet, it
wouldn't be able to identify and generate rules for attacks that
require detailed options, such as the HTTP PESC, RCE and LFI.
